---
title: "Online Assessment"
author: "Nsubuga Emmnauel Reagan"
date: "`r Sys.Date()`"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

DATA ANALYSIS ASSESSMENT

LOAD PACKAGES

```{r}
library(tidyverse)
library(scales)
library(readxl)
```

LOAD THE  CAMP DATASET

```{r}
workbook_path <- "R:/R Programming/Raising the village/Camp1.xlsx"
```

LOAD THE SHEETS

```{r}
# List all sheet names
sheet_names <- c('Treatment Baseline', 'Control Baseline', 'Treatment Midline', 'Control Midline', 'Treatment Endline', 'Control Endline')

```



```{r}
# Create a list to store data frames
sheets_list <- list()

# Read each sheet into a separate data frame and store in the list
for (sheet in sheet_names) {
  sheets_list[[sheet]] <- read_excel(workbook_path, sheet = sheet)
}

# Assign each data frame in the list to a variable in the global environment
list2env(sheets_list, .GlobalEnv)
```
PRINTING THE DATA FRAMES
```{r}
print(head(`Treatment Baseline`))
```

CONTROL BASELINE

```{r}
print(head(`Control Baseline`))
```

TREATMENT MIDLINE

```{r}
print(head(`Treatment Midline`))
```

CONTROL MIDLINE

```{r}
print(head(`Treatment Midline`))
```

TREATMENT ENDLINE

```{r}
print(head(`Treatment Midline`))
```

CONTROL ENDLINE

```{r}
print(head(`Control Endline`))
```


ASSIGNMENT 1: CONVERSION TO USD


```{r}
# Define conversion rates
conversion_rates <- list(
  Baseline = 3400,
  Midline = 3600,
  Endline = 4000
)

```

DEFINE THE COLUMNS TO CONVERT:

```{r}
columns_to_convert <- c(
  "Formal Employment (Ugx)",
  "Personal Business & Self Employment (Ugx)",
  "Casual Labour (Ugx)",
  "Remittances & Gifts (Ugx)",
  "Rent Income (Property & Land) Ugx",
  "Seasonal Crops Income (Ugx)",
  "Perenial Crops Income (Ugx)",
  "Livestock Income (Ugx)",
  "HH Income"
)

```


```{r}
library(stringr)

# Function to convert UGX to USD and rename columns
convert_to_usd <- function(df, rate) {
  existing_columns <- columns_to_convert[columns_to_convert %in% colnames(df)]
  
  if (length(existing_columns) > 0) {
    df <- df %>%
      mutate(across(all_of(existing_columns), ~ . / rate)) %>%
      rename_with(~ str_replace(., "\\(Ugx\\)", "(USD)"), all_of(existing_columns))
  }
  
  return(df)
}
```


APPLY CONVERSION RATE

```{r}
# Apply conversion rates and rename columns for each data frame
for (sheet in names(sheets_list)) {
  if (grepl("Baseline", sheet, ignore.case = TRUE)) {
    sheets_list[[sheet]] <- convert_to_usd(sheets_list[[sheet]], conversion_rates$Baseline)
  } else if (grepl("Midline", sheet, ignore.case = TRUE)) {
    sheets_list[[sheet]] <- convert_to_usd(sheets_list[[sheet]], conversion_rates$Midline)
  } else if (grepl("Endline", sheet, ignore.case = TRUE)) {
    sheets_list[[sheet]] <- convert_to_usd(sheets_list[[sheet]], conversion_rates$Endline)
  }
}

# Assign each data frame in the list to a variable in the global environment
list2env(sheets_list, .GlobalEnv)
```

ASSIGNMENT 1 QUESTION 2

DATA PREPARATION

```{r}
# Combine data for Baseline, Midline, and Endline
baseline_data <- bind_rows(`Treatment Baseline` %>% mutate(group = "treatment", time = "baseline"),
                           `Control Baseline` %>% mutate(group = "control", time = "baseline"))

midline_data <- bind_rows(`Treatment Midline` %>% mutate(group = "treatment", time = "midline"),
                          `Control Midline` %>% mutate(group = "control", time = "midline"))

endline_data <- bind_rows(`Treatment Endline` %>% mutate(group = "treatment", time = "endline"),
                          `Control Endline` %>% mutate(group = "control", time = "endline"))

# Combine all data into one data frame
combined_data <- bind_rows(baseline_data, midline_data, endline_data)

```


```{r}
# Ensure correct types for group and time variables
combined_data <- combined_data %>%
  mutate(group = factor(group, levels = c("control", "treatment")),
         time = factor(time, levels = c("baseline", "midline", "endline")))
```


RUNNING THE DiD REGRESSION

```{r}
# Load necessary packages
install.packages("lmtest")
library(lmtest)


```


DEFINE THE MODEL

```{r}
# Define the model formula for each variable
model_formulas <- list(
  PPI = "`PPI(%)` ~ group * time",
  Livestock_Asset_Value = "`Livestock Asset Value` ~ group * time",
  HH_Income = "`HH Income` ~ group * time",
  HH_Income_Consumption_Residues_Per_Day = "`HH Income + Consumption + Residues` ~ group * time"
)
```

RUN THE REGRESSION

```{r}
# Run the DiD regression for each variable
results <- lapply(model_formulas, function(formula) {
  lm(as.formula(formula), data = combined_data)
})
```

THE RESULTS

```{r}
# Display the summary of each model
lapply(results, summary)
```

INTREPRETATION OF THE RESULTS

# PPI (%)

Intercept: The average PPI (%) for the control group at the baseline is 32.70.

Group treatment: The difference between the treatment and control groups at baseline is -1.09, 
which is not statistically significant (p-value = 0.38401).

time midline: The change in PPI for the control group from baseline to midline is -5.84, which is statistically significant (p-value = 6.56e-08).

time endline: The change in PPI for the control group from baseline to endline is -6.81, which is statistically significant (p-value = 2.72e-10).

group treatment: The interaction term is -0.83, indicating the difference-in-differences estimate at midline, which is not statistically significant (p-value = 0.59740).

group treatment: The interaction term is -4.59, indicating the difference-in-differences estimate at endline, which is statistically significant (p-value = 0.00359).


# Livestock Asset Value

Intercept: The average Livestock Asset Value for the control group at the baseline is 623634.

group treatment: The difference between the treatment and control groups at baseline is -27333,
which is not statistically significant (p-value = 0.844).

time midline: The change in Livestock Asset Value for the control group from baseline to midline is -27146, which is not statistically significant (p-value = 0.820).

time endline: The change in Livestock Asset Value for the control group from baseline to endline is -96694, which is not statistically significant (p-value = 0.416).

group treatment: The interaction term is 682186, indicating the difference-in-differences estimate at midline, which is statistically significant (p-value = 8.06e-05).

group treatment: The interaction term is 951909, indicating the difference-in-differences estimate at endline, which is statistically significant (p-value = 4.76e-08).

# HH Income

Intercept: The average HH Income for the control group at the baseline is 318.69.

group treatment: The difference between the treatment and control groups at baseline is 41.98, which is not statistically significant (p-value = 0.9165).

time midline: The change in HH Income for the control group from baseline to midline is 194.65, which is not statistically significant (p-value = 0.5711).

time endline: The change in HH Income for the control group from baseline to endline is 269.43, which is not statistically significant (p-value = 0.4317).

group treatment: The interaction term is 65.02, indicating the difference-in-differences estimate at midline, which is not statistically significant (p-value = 0.8961).

group treatment: The interaction term is 1118.68, indicating the difference-in-differences estimate at endline, which is statistically significant (p-value = 0.0257).


# HH Income + Consumption + Residues/Day

Intercept: The average HH Income + Consumption + Residues/Day for the control group at the baseline is 1647069.

group treatment: The difference between the treatment and control groups at baseline is 185502,
which is not statistically significant (p-value = 0.84958).

time midline: The change in HH Income + Consumption + Residues/Day for the control group from baseline to midline is 1493744, which is marginally significant (p-value = 0.07515).

time endline: The change in HH Income + Consumption + Residues/Day for the control group from baseline to endline is 2051368, which is statistically significant (p-value = 0.01427).

group treatment: The interaction term is 535965, indicating the difference-in-differences estimate at midline, which is not statistically significant (p-value = 0.65947).

grouptreatment: The interaction term is 3257865, indicating the difference-in-differences estimate at endline, which is statistically significant (p-value = 0.00782).


## QUESTION 3

To determine if RTV achieved the targets of $1.25/day at Midline and $2/day at Endline for HH Income, I need to calculate the mean HH Income for the treatment group at Midline and Endline,
Compare these means to the targets and visualize the results.

COMPUTING THE MEAN

```{r}
# Compute mean HH Income per day for the treatment group at Midline and Endline
mean_income_per_day <- combined_data %>%
  filter(group == "treatment") %>%
  group_by(time) %>%
  summarize(mean_HH_Income_per_day = mean(`HH Income`, na.rm = TRUE) / 30)  # Assuming monthly data with 30 days

# Display the means
print(mean_income_per_day)


```

COMPARISON TO TARGETS

```{r}
# Define the targets per day
targets_per_day <- data.frame(
  time = c("midline", "endline"),
  target = c(1.25, 2.00)
)

# Merge the means with the targets
comparison_per_day <- merge(mean_income_per_day, targets_per_day, by = "time")

# Check if targets were achieved
comparison_per_day <- comparison_per_day %>%
  mutate(achieved = ifelse(mean_HH_Income_per_day >= target, "Yes", "No"))

# Display the comparison
print(comparison_per_day)

```

VISUALISING THE RESULTS

```{r}
# Load necessary library for visualization

library(ggplot2)

# Plot the results
ggplot(comparison_per_day, aes(x = time, y = mean_HH_Income_per_day)) +
  geom_bar(stat = "identity", fill = "skyblue", width = 0.5) +
  geom_hline(aes(yintercept = target, color = time), linetype = "dashed") +
  geom_text(aes(label = paste("Achieved:", achieved), y = mean_HH_Income_per_day + 0.1), vjust = 0) +
  labs(title = "Mean HH Income per Day at Midline and Endline",
       x = "Time",
       y = "Mean HH Income per Day (USD)") +
  theme_minimal() +
  scale_color_manual(values = c("midline" = "red", "endline" = "blue")) +
  theme(legend.position = "none")

```

ASSIGNMENT 1: QUESTION 4

CONVERT RENT INCOME TO USD

```{r}
# Function to convert and rename "Rent Income (Property & Land) Ugx" to USD
convert_rent_income_to_usd <- function(df, rate) {
  if ("Rent Income (Property & Land) Ugx" %in% colnames(df)) {
    df <- df %>%
      mutate(`Rent Income (Property & Land) (USD)` = `Rent Income (Property & Land) Ugx` / rate) %>%
      select(-`Rent Income (Property & Land) Ugx`)
  }
  return(df)
}
```


```{r}
# Perform multiple linear regression
model <- lm(`HH Income` ~ `Formal Employment (USD)` + `Personal Business & Self Employment (USD)` + `Casual Labour (USD)` + `Remittances & Gifts (USD)` + `Seasonal Crops Income (USD)` + `Perenial Crops Income (USD)` + `Livestock Income (USD)`, data = combined_data)

# Display the summary of the regression model
summary(model)

```

All the variables significantly contribute to HH Income, but the income from Perenial Crops is the most significant contributor.

# REASONING

I've used multiple linear regression because it allows us to understand the relationship between a dependent variable (HH Income) and multiple independent variables (different sources of income). The coefficients from the regression model help identify the contribution of each independent variable to the dependent variable.


# ASSIGNMENT TWO

LOADING THE WORKBOOK INTO MY ENVIRONMENT

```{r}
# Define the path to the uploaded workbook
workbook_path <- "R:/R Programming/Raising the village/Base 1.xlsx"

# List all sheet names
sheet_names <- excel_sheets(workbook_path)

# Read each sheet into a separate data frame and store them in a list
sheets_list <- lapply(sheet_names, function(sheet) {
  read_excel(workbook_path, sheet = sheet)
})

# Combine the sheets into one data frame
combined_data <- bind_rows(sheets_list)

# Display the first few rows of the combined data
print(head(combined_data))

```

SUMMARY STATISTICS

```{r}
summary(combined_data)
```


The dataset comprises 2513 records, offering a detailed look at both demographic and economic variables. The average age is around 50 years, ranging from 18 to 103, with a slight male bias. Education levels average around six years, though some unusual negative values suggest a need for data cleaning. About 65% of the individuals are literate, and the average household size is approximately 4.5 members.

Economically, there's a broad spectrum of income sources. The average household income is about 4.24 million Ugx, but there's significant variability, with some households earning extremely high incomes, especially from perennial crops. This points to considerable economic disparity.

A significant portion of the data is missing, with around 1263 missing values in many columns, roughly half the dataset. This will require addressing, either through imputation or by excluding incomplete records, for accurate analysis.

Income sources like formal employment, personal business, and casual labor vary widely, with many households reporting zero income from these sources, indicating reliance on other streams.

# QUESTION 2: ASSIGNMENT 2

```{r}
# Calculate the 1st and 94th percentiles for HH Income
lower_bound <- quantile(combined_data$`HH Income`, 0.01, na.rm = TRUE)
upper_bound <- quantile(combined_data$`HH Income`, 0.94, na.rm = TRUE)

# Remove outliers
cleaned_data <- combined_data %>%
  filter(`HH Income` >= lower_bound & `HH Income` <= upper_bound)

# Display the first few rows of the cleaned data
print(head(cleaned_data))

```

QUESTION 3: ADJUSTING FOR INFLATION

Adjusting for local inflation is essential to ensure that the values of economic variables reflect their true purchasing power over time. Inflation erodes the value of money, meaning that the same amount of currency can buy fewer goods and services in the future compared to today. 
By adjusting for inflation, we can compare values from different years on a like-for-like basis, maintaining the real value of money.

DEFINING THE INFLATION RATES

```{r}
# Define the inflation rates for each year relative to 2017
inflation_rates <- c(
  "2017" = 1,
  "2018" = 1.03,
  "2019" = 1.04,
  "2020" = 1.05,
  "2021" = 1.06
)

```


VARIABLES

```{r}
# Variables to adjust for inflation
vars_to_adjust <- c(
  "Formal Employment (Ugx)", "Personal Business & Self Employment (Ugx)", 
  "Casual Labour (Ugx)", "Season 1 Agriculture Value (Ugx)", 
  "Season 2 Agriculture Value (Ugx)", "Perennial Agriculture Value (Ugx)",
  "Livestock Income / Consumed (Ugx)", "HH Income", 
  "HH Income + Consumption + Residues", "HH Income + Consumption+ Assets + Residues"
)
```

