---
title: "Ensemble Prediction - Decision Tree & Neural Network"
author: "Sheik Mohamed Imran"
date: "December 4, 2017"
output: html_document
---


###Load the required Libraries
```{r, message=F, warning=F}
library(funModeling)
library(caret)
library(nnet)
library(gmodels)
library(C50)
library(NeuralNetTools)
library(partykit)
library(knitr)
knitr::opts_chunk$set(fig.width=12, fig.height=12) 
library(e1071)   
```

###Read the data into the variable
```{r}
cancerdata<-read.csv("../input/data.csv")
```

###Check dataset for NA Data

```{r}
sum(complete.cases(cancerdata)) == nrow(cancerdata)
```

Here, 'FALSE' indicates 'NA' are available, now lets identify the corresponding column(s).

###Identify columns with NA
```{r}
colnames(cancerdata)[colSums(is.na(cancerdata)) > 0]
```

Column 'X' is identified as the column containing only 'NA'

###Remove unwanted columns from the original data
'X' has all the values as 'NA' and from the description of the dataset, 'id' corresponds to the patient identifier. These columns can be removed to have a sane dataset.
```{r}
cancerdata <- subset(cancerdata, select = -c(X))
cancerdata <- subset(cancerdata, select = -c(id))
cancerdata$diagnosis<-as.factor(cancerdata$diagnosis) 
```

###Basic Analysis

```{r, results=F}
cancerdata_status=df_status(cancerdata)
```

```{r}
kable(cancerdata_status, caption = "Breast Cancer Data Analysis")
```

As one can notice, there are some column containing values as '0', but not any 'NA' or 'INF' values, this data is clean enough to proceed with further processing

###Plot the data
```{r}
plot_num(cancerdata)
```

In most of the columns, data frequency is normally distributed

###Normailize dataset using PCA
Column 1 is the diagonised value (M=malignant and B=benign), we will ignore this to conduct PCA and prediction
```{r}
pca <- prcomp(cancerdata[,-1], retx=TRUE, center=TRUE, scale=TRUE)
pred <- predict(pca, newdata=cancerdata[,-1])
```


###Include the removed column back to the standardised data
```{r}
cancerdata <- data.frame(cbind(pred,diagnosis=cancerdata[,1]))
cancerdata$diagnosis=as.factor(cancerdata$diagnosis)
```


###Create sample and train data
Here we create the sample and train data using seed function. This value is set to a predefined value such that the same set of rows are returned throught the code.
```{r}
set.seed(1023)
samp <- sample(nrow(cancerdata), nrow(cancerdata)*0.8)
cancerdata.train <- data.frame(cancerdata[samp,])
cancerdata.valid <- data.frame(cancerdata[-samp,])
```

##Create C5.0 Model 
We now create a basic C5.0 model
```{r}
model_c5<-C5.0(diagnosis ~ ., data=cancerdata.train,
           trials=50,
           control = C5.0Control
           (                        
             noGlobalPruning = T,
             CF=0.8,
             minCases=10,
             sample = 0.80,
             winnow=F,
             earlyStopping=T
           ))
```

###List important variables in the model
```{r}
C5imp(model_c5)
```

Based on the output, we can see that the most important node variable is PC1, followed by PC2 and PC5

###Plot the created model
The default plot function for the C5.0, gives a greyscale image with features. We have extended the partykit function to plot a colourful descion tree of our C5.0 model.
```{r}
columncol<-hcl(c(270, 260, 250), 200, 30, 0.6)
labelcol<-hcl(200, 200, 50, 0.2)
indexcol<-hcl(150, 200, 50, 0.4)

model_c5a=C50:::as.party.C5.0(model_c5)
plot(model_c5a,type="simple",gp = gpar(fontsize = 8), drop_terminal = TRUE, tnex=1,
     inner_panel = node_inner(model_c5a, abbreviate = FALSE,
                              fill = c(labelcol, indexcol), pval = TRUE, id = TRUE),
     terminal_panel=node_barplot(model_c5a, col = "black", fill = columncol[c(1,2,4)], 
                                 beside = TRUE, ymax = 1, ylines = TRUE, widths = 1, gap = 0.1,
                                 reverse = FALSE, id = TRUE))

```




###Make predictions and print the crosstable
Create prediction of validation data based on the C5.0 model created and print the results in a CrossTable.
```{r}
predictions_c5 <- predict(model_c5, cancerdata.valid[-31], type="class")
CrossTable(cancerdata.valid$diagnosis, predictions_c5, prop.chisq = FALSE,
           prop.c = FALSE, prop.r = FALSE, dnn = c("Actual diagnosis",
                                                   "Predicted diagnosis"))

```

###Check the accuracy of the model
```{r}
paste0(round(sum(predictions_c5 == cancerdata.valid$diagnosis)/nrow(cancerdata.valid)*100,digit=4),"%")
```

###Add predictions from C5.0 to train data, for better accuracy in subsequent model
```{r}
cancerdata.train$predictions_c5 <- predict(model_c5, cancerdata.train[-31], type="class")
cancerdata.train$predictions_c5=as.factor(cancerdata.train$predictions_c5)
```

##Create Neural network model
```{r,results=F}
myform <- as.formula(paste0('diagnosis ~ ', 
                            paste(names(cancerdata.train[!names(cancerdata.train) %in% 'diagnosis']),
                                  collapse = ' + ')))
model_nnet<-nnet(diagnosis ~. , data=cancerdata.train,size=3,rang = 1,decay = 8e-4, maxit = 200)
```

###Make predictions and print the crosstable
Create prediction of validation data based on the Neural Net model created and print the results in a CrossTable.
```{r}
predictions_nnet<-predict(model_nnet, cancerdata.valid[,-31], type = c("class") )
CrossTable(cancerdata.valid$diagnosis, predictions_nnet, prop.chisq = FALSE,
           prop.c = FALSE, prop.r = FALSE, dnn = c("Actual diagnosis",
                                                   "Predicted diagnosis"))

```

##Print the accuracy of the model
```{r}
paste0(round(sum(predictions_nnet == cancerdata.valid$diagnosis)/nrow(cancerdata.valid)*100,digit=4),"%")
```


###Evaluate variable importance
Relative importance of input variables in neural networks as the sum of the product of raw inputhidden, hidden-output connection weights, proposed by Olden et al. 2004.

```{r}
olden(model_nnet)
```

In the plot, here are both positive and negative values.The values reflect negative and positive relationships between the variable and the response variable.


###Plot the neuralnetwork
Here we specifiy the relative importance of input variables in neural networks using Garson’s algorithm. The variables of importance are highlighted in green, darker being higher importance.

```{r}
rel_imp <- garson(model_nnet, bar_plot = FALSE)$rel_imp
cols <- colorRampPalette(c('lightgreen', 'darkgreen'))(3)[rank(rel_imp)]
plotnet(model_nnet, cex_val = .8,max_sp=TRUE,circle_cex=3,circle_col = list(cols, 'lightblue'))
```

####References:
- https://amunategui.github.io/blending-models/
- https://rmanic.wordpress.com/2016/05/19/creating-a-well-formatted-decision-tree-with-partykit-and-listing-the-rules-of-the-nodes/
- https://beckmw.wordpress.com/2013/11/14/visualizing-neural-networks-in-r-update/