---
title: "EDA"
author: "Michael Griffiths"
date: "December 29, 2016"
output: html_document
---

```{r}
library(needs)
needs(tidyverse,
      plotly,
      rpart, 
      rpart.plot,
      viridis,
      Metrics,
      glmnet,
      formattable)
```

We have some data out of a HR system, related to employees leaving the company. Let's take a look!

# Data

```{r}
data <- read_csv("../input/HR_comma_sep.csv")

data %>% 
  head(5e2) %>% 
  formattable() %>% 
  as.datatable
```

We can make a few observations about the columns --

  * `satisfaction_level`: A numeric indictor, presumably (?) filled out by the employee.
  * `last_evaluation`: A numeric indicator, presumably filled in by the employee's manager.
  * `number_project`: A integer - perhaps the number of projects the employee has worked on. Is this something like construction?
  * `average_monthly_hours`: The number of hours they work (billed?) in the month. On average?
  * `time_spend_company`: An integer value, perhaps years of service.
  * `Work_accident`: Looks like a boolean value, probably whether or not they had an accident.
  * `left`: Looks like a boolean value, leave or not.
  * `promoted_last_5years`: Looks like a boolean value.
  * `sales`: Not sure what this means. Part of the sales force? 
  * `salary`: Looks like a 3-level salary score (`low`, `medium`, `high`)
  
Some missing things - 

  * `user_id`: What if an employee leaves and comes back? Are they in there twice?
  
  
Well, let's investigate some of the columns.

## Salary

What's the distribution of salary?

```{r}
data %>% 
  count(salary) %>% 
  formattable(align = 'l')
```

Unsurprisingly, only a small number of people have a `high` salary level. It's interesting that `medium` and `low` are around the same size.

## sales

```{r}
data %>% 
  count(sales) %>% 
  formattable(align = 'l')
```

It looks like we have people from all departments! That's interesting -- I wouldn't think `number_projects` would really apply to, say, **IT** or **accounting**.

It also looks like **sales** is the largest category, which makes sense. 

How does `sales` and `salary` interact?

```{r}
data %>% 
  count(sales, salary) %>% 
  mutate(
    salary = ordered(salary, c("low", "medium", "high")),
    n = comma(n, 0)
  ) %>% 
  spread(salary, n) %>% 
  formattable(list(area(T, 2:4) ~ color_tile("white", "orange")), align = 'l')
```

Some pretty obvious stuff here.

  * **management** tends to have higher salaries.
  * Very few people in **sales** have high salaries.
  * There are _a lot_ of people with lower salaries in **support**, **sales**, and **technical**.
  
Let's look at the same chart by percentage instead of actual numbers.

```{r}
data %>% 
  count(sales, salary) %>% 
  group_by(sales) %>% 
  mutate(
    salary = ordered(salary, c("low", "medium", "high")),
    n = percent(n, 0),
    n = n / sum(n)
  ) %>% 
  spread(salary, n) %>% 
  formattable(list(area(T, 2:4) ~ color_tile("white", "orange")), align = 'l')
```

That makes the distribution of **management** really obviousl, and also points out that the second highest is **accounting** (which was not obvious to me in the precious chart). 

It also seems that while there are _a lot_ of people in some of the other departments with low/medium salaries; as a percentage of each they're all roughly the same. If we wanted to make it really obvious, we could index. But given how small the variation, there's no real need.

## Promoted

Now let's look at `promotion_last_5years`.

```{r}
data %>% 
  count(promotion_last_5years) %>% 
  formattable(align = 'l')
```

The short answer is "not many" people were promoted. 

```{r}
data %>% 
  count(promotion_last_5years, salary) %>% 
  group_by(salary) %>% 
  mutate(n = n / sum(n), n = percent(n)) %>% 
  spread(promotion_last_5years, n) %>% 
  formattable(align = 'l')
```

It's unsurprising to note that, of the people making a high salary, nearly 6% of them were promoted in the last 5 years.

## Time Spent

```{r}
data %>% 
  count(time_spend_company) %>% 
  ggplot(aes(time_spend_company, n)) +
  geom_line() ->
  g

ggplotly(g, width = 900)
```

I always find the spikes interesting - a large one at three years, and then a serious dropoff for year four onward.

It's interesting that they're not counting people who have been there for less than two years. What's that about? Is there too much churn?

How does that relate to salary?

```{r}
data %>% 
  mutate(salary = ordered(salary, c("low", "medium", "high"))) %>% 
  count(time_spend_company, salary) %>% 
  group_by(time_spend_company) %>% 
  mutate(n = percent(n / sum(n))) %>% 
  ggplot(aes(time_spend_company, n)) +
  geom_area(aes(fill = salary)) ->
  g

ggplotly(g, width = 900)
```

Ideally, I'd like to make a story like "salary is pretty flat until around 5 or 6 years, and then the people who've been here longer tend to make more money." But it doesn't look like that -- note the spike in **low** salary levels at ~8 years. 

Still, the _ratio_ of people with different salary levels follows roughly that disribution.

Note that there are a decent number of **high** salary people with < 3 years of service. I'd wager there's a group of managers/specialists who are highered at the senior level, then only stick around for a few years. The people who remain might have fewer options, which would explain why there are fewer **high** salary people at year 5 or so.

## Monthly Hours

Let's take a look at monthly hours.

```{r}
data %>% 
  ggplot(aes(average_montly_hours)) +
  geom_density(fill = "orange") ->
  g

ggplotly(g, width = 900)
```

Interesting that we have what appears to be a bimodal distribution - one peak at ~150 hours (roughly a 40 hour work week), and one at 260 (a crazy-insane 65 hour work-week).

How does that break down by department and by salary level?

```{r}
data %>% 
  mutate(salary = ordered(salary, c("low", "medium", "high"))) %>% 
  ggplot(aes(average_montly_hours)) +
  geom_density(fill = "orange") +
  facet_grid(sales ~ salary) -> 
  g

ggplotly(g, height = 1e3, width = 900)
```

I'm honestly surprised to not see more variation. However, there are a few interesting things we can point out.

  * Nearly all the distributions remain bimodal. Is there some other factor going on?
  * In **marketing** it seems like high-salary people take a break. It's more common to be working fewer hours.
  * In **support**, working longer is apparently required.
  * It seems like **hr** people at a medium salary don't work as much as high or low salary people.
  
## Project Counts

Let's see how many projects people work on.

```{r}
data %>% 
  count(number_project, sales) %>% 
  spread(number_project, n) %>% 
  formattable(align = 'l')
```

I'm honestly not sure what this variable - number of projects - even means. Does it mean e.g. clients they touch? Or is it "project ID" -- there are six projects? That's feasible, I suppose. 

Let's look at the percentage breakdown.

```{r}
data %>% 
  count(number_project, sales) %>% 
  group_by(number_project) %>% 
  mutate(n = percent(n / sum(n), 0)) %>% 
  spread(number_project, n) %>% 
  formattable(align = 'l')
```

Hmm -- it's reletively similar across projects. Does that suggest it's not "number of projects" but rather "ID of project they're assigned to?" Not necessarily. But why is there no `number_project` == 1?

The documentation says it's _Number of projects_, so we'll leave it there for now.

```{r}
data %>% 
  count(number_project, left) %>% 
  spread(left, n) %>% 
  formattable(align = 'l')
```

It's kind of suspicious that people with `number_project == 7` have all left the company...

Oh well.

## Satisfaction & Rating

```{r}
data %>% 
  mutate(left = factor(left, labels = c("Remain", "Left"))) %>% 
  ggplot(aes(satisfaction_level)) +
  geom_density(fill = "orange") +
  facet_wrap( ~ left, ncol = 2) -> 
  g

ggplotly(g, width = 900)
```

Well, that's pretty clear. People with low satisfaction rarely stay at the company - and I'd eye the people with satisfaction < 0.5 as ripe to leave - and those who leave fall into three categories:

  1. The "haters" with satisfaction < 0.2
  2. The "unsatisfied" with satisfaction < 0.5
  3. The "pretty good", with satisfaction > 0.6
  
How about rating?

```{r}
data %>% 
  mutate(left = factor(left, labels = c("Remain", "Left"))) %>% 
  ggplot(aes(last_evaluation)) +
  geom_density(fill = "orange") +
  facet_wrap( ~ left, ncol = 2) -> 
  g

ggplotly(g, width = 900)
```

I mean, that's pretty amazing. It suggests that the people who leave are either:

  1. Pretty bad at their job. Note that the low end of the scale is 0.4.
  2. Excellent at their job, probably leaving for a promotion.
  
Can we look at both satisfaction and last evaluation?

```{r, fig.width = 10, fig.height = 6}
data %>% 
  mutate(left = factor(left, labels = c("Remain", "Left"))) %>% 
  ggplot(aes(last_evaluation, satisfaction_level)) +
  geom_point(alpha = 0.05, color = "orange") +
  facet_wrap( ~ left, ncol = 2) -> 
  g

# Use WebGL for speed
ggplotly(g) %>% 
  toWebGL()
```

Well, hot damn: that's a pretty tight distribution for **Left**. Crazy tight.

# Leaving

Now let's look at why people left. Since we're interested in _understanding_, we can accomplish this a couple of ways - linear models or decision trees. Based on some of the other analyses performed it looks like random forests wins.

So let's try a single decision tree!

First we'll separate into test and train.

```{r}
n <- nrow(data)
idx <- sample(n, n * .66)

# Make a few modications
data %>% 
  mutate(
    left = factor(left, labels = c("Remain", "Left")),
    salary = ordered(salary, c("low", "medium", "high"))
  ) -> 
  d

train <- d[idx, ]
test <- d[-idx, ]
```

Then we'll train a single decision tree using `rpart` to and evaluate to see how good our fit is.

```{r}
tree <- rpart(left ~ ., data = train)

res <- predict(tree, test)

auc(as.numeric(test$left) - 1, res[, 2])
```

I mean, that's a pretty unbeliebable AUC score for a single tree. Let's take a look at it.

```{r, fig.width = 10, fig.height = 8}
rpart.plot(tree, type = 2, fallen.leaves = F, cex = 1, extra = 2)
```

So, what can we observe?

  1. Satisfaction level appears to be the most import piece. If you're above 0.46 you're much more likely to stay (which is what we observed above).
  2. If you have low satisfaction, the number of projects becomes import. If you're on more projects you're more likely to remain. If you're on fewer projects -- perhaps you see the writing on the wall?
  3. If you're happy, have been at the company for less than 4.5 years, and score over 81% on your last evaluation, you're very likely to leave. And, it appears as if the "decider" is monthly hours over 216. 
  
In brief:

  1. If you're successful and overworked, you leave.
  2. If you're unhappy and overworked, you leave.
  3. If you're unhappy and underworked, you leave. 
  4. If you've been at the company for more than 6.5 years, you're more likely to be happy working longer hours.
  
Let's limit the variables that go into the model to `satisfaction_level`, `last_evaluatoin`, and `average_monthly_hours`.

```{r}
tree <- rpart(left ~ satisfaction_level + last_evaluation + average_montly_hours, data = train)

res <- predict(tree, test)

auc(as.numeric(test$left) - 1, res[, 2])
```

Still an excellent AUC score.

```{r, fig.width = 10, fig.height = 8}
rpart.plot(tree, type = 2, fallen.leaves = F, cex = 1, extra = 2)
```

We see much the same story -- if you're good and overworked you leave; if you're unhappy you tend to leave, especially if you're not getting enough work. 

# Promotion

How about people getting a promotion?

Let's reprise our data splitting.

```{r}
n <- nrow(data)
idx <- sample(n, n * .66)

# Make a few modications
data %>% 
  mutate(
    left = factor(left, labels = c("Remain", "Left")),
    salary = ordered(salary, c("low", "medium", "high")),
    promotion_last_5years = factor(promotion_last_5years, labels = c("Not Promoted", "Promoted"))
  ) -> 
  d

train <- d[idx, ]
test <- d[-idx, ]
```

Now we can train and examine.

```{r}
tree <- rpart(promotion_last_5years ~ ., data = train)

res <- predict(tree, test)

auc(as.numeric(test$promotion_last_5years) - 1, res[, 2])
```

Huh: that's awful. Do we really have such limited ability to predict a promotion?

```{r, fig.width = 10, fig.height = 6}
rpart.plot(tree, type = 2, fallen.leaves = F, cex = 1, extra = 2)
```

So... this suggests that the best measure of "having been promoted" is "been at the company for a while". Interesting. 

However, with such a low AUC score it's obviously untrustworthy. 

We can try to see what features matter with a lasso-penalized logistic regression. Let's take a quick look.

```{r}
X = model.matrix(promotion_last_5years ~ 0 + ., data) 
y = data$promotion_last_5years

lasso <- glmnet(X, y)
```

All we're really interested in is the coefficient chart (from e.g. `plot(lasso)`). But I want the coefficient label and value to be available on hover, so alas we must deign to manipulate the `glmnet` object directly.

```{r}
c = as.data.frame(as.matrix(coef(lasso)))
c %>% 
  mutate(var = rownames(c)) %>% 
  gather(lambda, estimate, -var) %>% 
  filter(var != "(Intercept)") %>% 
  group_by(lambda) %>% 
  mutate(l1_norm = sum(abs(estimate))) %>% 
  ggplot(aes(l1_norm, estimate)) +
  geom_line(aes(color = var)) +
  geom_hline(yintercept = 0) +
  theme(legend.position = "null") ->
  g

ggplotly(g, width = 900) %>%
  layout(showlegend = FALSE)
```

So, in the case of the linear model, we get a slightly different conclusion. 

  1. The best indicator of "having been promoted" is "being in management". Shocking - I know - though it makes sense if people are promoted _from_ an IC role _into_ management. Therefore, promotions could primarily indicate management responsibility.
  2. The next best indicator is "salary low". It makes sense that if you're underpaid you might get a promotion.
  3. Then it's a whole lot of things that matter. Unsurprisingly, if you leave the company you can't get promoted.
  
# Conclusion

Overall, I'd say that:

  * Figuring out why people are leaving is pretty easy.
  * Identifying promotions seems very hard.
  * Having salary cut into "high", "medium", and "low" is not something I really like. 
  
Also, I've really enjoyed using `plotly` for charting here. The conversion from `ggplot2` is shockingly good. Plus, the `toWebGL` method (which I only just leared about) makes it feasilble to plot thousands of points without horribad performance.
